### 2019-01-15 writing/thesis/chapters/conclusion.tex
Update everything.

\chapter{Conclusion}
Conclusion here.

>>>>

\chapter{Conclusion}\label{chap:conclusion}
We implemented a \muscl{}-limited \aderdg{}-scheme with \amr{} for the reactive Navier Stokes equations in this thesis.
We showed that the scheme works and is capable of simulation different scenarios:\todo{Reference sections of results. Reference theory chapter. Reference implementation.}
\begin{itemize}
\item We showed in \cref{sec:results-convergence} that our scheme converges for all considered orders for the Navier-Stokes equations.
\item We evaluated our method for standard \textsc{cfd} scenarios and achieved competitive results for all considered scenarios.
  This is true for two-dimensional scenarios (Taylor-Green, Lid-Driven Cavity) and for the three-dimensional \textsc{abc}-flow.
\item Furthermore, our method allows us to simulate flows in hydrostatic equilibrium correctly.
\item The results of \cref{sec:results-tts-amr} showed that our \amr{}-criterion is able to save a significant portion of the computational effort compared to a uniform fine grid.
  It is able to do this without degrading the quality of the solution, even for a scenario that contains small-scale details.
\item Chemical reactions and limiter work for non-stiff detonations.
\end{itemize}

Of course, there are some things than can be improved and which need further work, such as:
\begin{itemize}
\item The current treatment of the boundary conditions is not stable for every scenarios.
  Especially the viscous walls for the \muscl{}-method are not correct, due to limitations of \exahype{}.
\item Due to the \textsc{cfl}-condition~\pcref{eq:cfl-aderdg,eq:cfl-muscl}, the timestep of our \aderdg{}-scheme scales quadratically with the number of grid cells.
  Even though our \amr{}-criterion is able to reduce the number of needed cells drastically, we could get another large improvement by utilizing local timestepping.
\item Our implementation of the \textsc{cj}-detonation wave cannot handle realistic detonations because we do not have an implementation of stiff source terms.
  An implementation of this is straight-forward but includes changes to the space-time predictor and its initial guess, and also to the \muscl{}-scheme.
\end{itemize}

### 2019-01-15 writing/thesis/chapters/introduction.tex
Update everything.

\chapter{Introduction}\label{chap:introduction}
We make the following contributions:
\begin{itemize}
\item In \cref{sec:navier-stokes}, we describe a coupling of the compressible Navier-Stokes equations with an advection-diffusion-reaction term for two dimensional problems.
\item We describe an \textsc{ader-dg} (\cref{sec:ader-dg}) and \textsc{muscl}-scheme (\cref{sec:muscl}) for problems with diffusive terms.
  In addition, we describe a combination of both methods in \cref{sec:limiting}, where we use the \textsc{muscl}-scheme in cases where the \textsc{ader-dg}-method is unstable.
\item In \cref{sec:amr}, we introduce a mesh refinement methods that checks, whether the local cell is an outlier.
  As indicator function we use the total variation of the discrete solution and as outlier detection we use the mean and standard variance of this function.
  We evaluate the performance of this method using a time-to-solution test.
\item We investigate the ability of our scheme to simulate classical fluid mechanics test scenarios correctly.
  For this we use a convergence test with a manufactured solution, the Taylor-Green vortex (2d) and the ABC-Flow (3d).
\item We describe the simulation of flows with a background state that is in hydrostatic equilibrium.
  This is useful for problems in atmospheric flow and for problems in meterology.
  Furthermore, we evaluate our discretization for these scenarios using two scenarios that simulate rising bubbles.
\item We show that our method is capable of simulating CJ-detonation waves that are a shock which originates from a chemical reaction.
\end{itemize}

>>>>

\chapter{Introduction}\label{chap:introduction}
\todo{General boilerplate introduction (short), mention exahype}

We make the following contributions:
\begin{itemize}
\item In \cref{sec:navier-stokes}, we describe a coupling of the compressible Navier-Stokes equations with an advection-diffusion-reaction term for two dimensional problems.
\item We describe an \aderdg{} (\cref{sec:ader-dg}) and \muscl{}-scheme (\cref{sec:muscl}) for problems with diffusive terms.
  In addition, we describe a combination of both methods in \cref{sec:limiting}, where we use the \muscl{}-scheme in cases where the \aderdg{}-method is unstable.
\item In \cref{sec:amr}, we introduce a mesh refinement methods that checks, whether the local cell is an outlier.
  As indicator function we use the total variation of the discrete solution and as outlier detection we use the mean and standard variance of this function.
  We evaluate the performance of this method using a time-to-solution test.
\item We investigate the ability of our scheme to simulate classical fluid mechanics test scenarios correctly.
  For this we use a convergence test with a manufactured solution, the Taylor-Green vortex ($2D$) and the \textsc{abc}-Flow ($3D$).
\item We describe the simulation of flows with a background state that is in hydrostatic equilibrium.
  This is useful for problems in atmospheric flow and for problems in meterology.
  Furthermore, we evaluate our discretization for these scenarios using two scenarios that simulate rising bubbles.
\item We evaluate the simulation of a \textsc{cj}-detonation wave that are a shock which originates from a chemical reaction.
\end{itemize}

### 2019-01-15 writing/thesis/chapters/methods.tex
Update everything.

% Variables and other equation stuff
\newcommand{\Q}{\bm{Q}}
\newcommand{\gradQ}{\gradient{\Q}}
\newcommand{\Qrho}{\rho}
\newcommand{\Qj}{\rho \bm{v}}
\newcommand{\Qv}{\bm{v}}
\newcommand{\QE}{\rho E}
\newcommand{\QZZ}{Z} % TODO: Name?
\newcommand{\QZ}{\rho \QZZ}
\newcommand{\potT}{\theta}
\newcommand{\backgroundPotT}{\overline{\theta}}
\newcommand{\pertubationPotT}{\theta'}
\newcommand{\stressT}{\bm{\sigma}}
\newcommand{\pressure}{p}
\newcommand{\maxConvEigen}[1][]{
  \vert%
  \lambda_c^{\text{max}}
  \notblank{#1}{\left(#1\right)}{}
  \vert%
}
\newcommand{\maxViscEigen}[1][]{
  \vert%
  \lambda_v^{\text{max}}
  \notblank{#1}{\left(#1\right)}{}
  \vert%
}
\newcommand{\Riemann}{\operatorname{Riemann}}

% Domain
\newcommand{\domain}{\Omega}
\newcommand{\boundary}{\partial \domain}
\newcommand{\broken}{\domain}

% Cells
\newcommand{\cell}[1][]{C_{#1}}
\newcommand{\refCell}[1][]{\hat{C}_{#1}}
\newcommand{\cellb}{\partial{} \cell}
\newcommand{\refCellb}{\partial{} \refCell}

% Basis functions
\newcommand{\lagrange}[1][i]{\varphi_{#1}}
\newcommand{\lagrangeRef}[1][i]{\hat{\varphi}_{#1}}
\NewDocumentCommand{\sbasis}{ O{} O{} }{\phi^{#1}_{#2}}
\NewDocumentCommand{\sbasisRef}{ O{} O{} }{\hat{\phi}^{#1}_{#2}} %{\hat{\sbasis[#1][#2]}}
\NewDocumentCommand{\stbasis}{ O{} O{} }{ \psi^{#1}_{#2}}
\NewDocumentCommand{\stbasisRef}{ O{} O{} }{\hat{\psi}^{#1}_{#2}} %{\hat{\stbasis[#1][#2]}}
%\newcommand{\stestfunction}[1]{\sbasis[#1]}
\NewDocumentCommand{\stestfunction}{ O{} O{} }{\sbasis[#1][#2]}
\NewDocumentCommand{\sttestfunction}{ O{} O{} }{\stbasis[#1][#2]}

% Space-time dofs
\NewDocumentCommand{\stpredictor}{ O{} O{} }{\overline{\bm{q}}^{#1}_{#2}}
\NewDocumentCommand{\stpredictorCoeff}{ O{} O{} }{\underline{\overline{q}}^{#1}_{#2}}

\NewDocumentCommand{\stflux}{ O{} O{} }{\overline{F}^{#1}_{#2}}
\NewDocumentCommand{\stfluxCoeff}{ O{} O{} }{\underline{\overline{F}}^{#1}_{#2}}

% \NewDocumentCommand{\stsol}{ O{} O{} }{\overline{u}^{#1}_{#2}}
% \NewDocumentCommand{\stsol}{ O{} O{} }{\underline{\overline{u}}^{#1}_{#2}}

\NewDocumentCommand{\stsource}{ O{} O{} }{\overline{S}^{#1}_{#2}}
\NewDocumentCommand{\stsourceCoeff}{ O{} O{} }{\underline{\overline{S}}^{#1}_{#2}}

% Space dofs
\NewDocumentCommand{\spredictor}{ O{} O{} }{\bm{q}^{#1}_{#2}}
\NewDocumentCommand{\spredictorCoeff}{ O{} O{} }{\underline{q}^{#1}_{#2}}

\NewDocumentCommand{\ssource}{ O{} O{} }{S^{#1}_{#2}}
\NewDocumentCommand{\ssourceCoeff}{ O{} O{} }{\underline{S}^{#1}_{#2}}

\NewDocumentCommand{\ssol}{ O{} O{} }{u^{#1}_{#2}}
\NewDocumentCommand{\ssolCoeff}{ O{} O{} }{\underline{u}^{#1}_{#2}}

\NewDocumentCommand{\sflux}{ O{} O{} }{F^{#1}_{#2}}
\NewDocumentCommand{\sfluxCoeff}{ O{} O{} }{\underline{F}^{#1}_{#2}}

% Equation parts
\newcommand{\flux}{F}
\newcommand{\viscFlux}{\flux^{v}}
\newcommand{\hyperFlux}{\flux^{h}}
%\newcommand{\source}{\bm{S}}
\newcommand{\source}[1][]{
  \notblank{#1}{
S_{#1}
}{
\bm{S}
}
}

% Integrals
\newcommand{\curTime}{t^k}
\newcommand{\nextTime}{t^{k+1}}
\newcommand{\intdt}[1]{\int_{\curTime}^{\nextTime} #1 \dd{t}}
\newcommand{\intdcell}[1]{\int_{\cell} #1 \dd{\bm{x}}}
\newcommand{\intdrefcell}[1]{\int_{\refCell} #1 \dd{\hat{\bm{x}}}}
\newcommand{\intdcellb}[1]{\int_{\cellb} #1 \dd{S}} % TODO: define boundary of cell
\newcommand{\intdrefcellb}[1]{\int_{\refCellb} #1 \hat{\dd{S}}} % TODO: define boundary of cell

% Yet to be sorted
\newcommand{\quadWeight}[1][i]{w_{#1}}
\newcommand{\quadNode}[1][i]{\hat{x}_{#1}}
\newcommand{\quadNodeNd}[1][i]{\hat{\bm{x}}_{#1}}
\newcommand{\sbasisSize}{(N+1)^\dimensions}
\newcommand{\stbasisSize}{(N+1)^{\dimensions+1}}
\newcommand{\normal}{\bm{n}}

\newcommand{\pleft}{\bm{L}}
\newcommand{\prightsol}{\bm{r}}
\newcommand{\prightpred}{\bm{w}}

@@

We conclude this chapter by a description of finite volume limiting and adaptive mesh refinement (\amr{}).
The limiting combines both numerical methods to achieve a stable, high-order method and the \amr{} allows us to perform high quality simulations with a smaller computational cost.

\section{Reactive Compressible Navier Stokes Equations}\label{sec:navier-stokes}
\newcommand{\diffCoeff}{\varepsilon}
\newcommand{\hyperFluxDef}{
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure) \\
    \Qj \QZZ
  \end{pmatrix}
}
\newcommand{\viscFluxDef}{
  \begin{pmatrix}
     -\diffCoeff \gradient{\Qrho}\\
     \stressT (\Q, \gradQ)  \\
     \Qv \cdot \stressT (\Q, \gradQ) - \kappa \gradient{T}\\
     -\diffCoeff \gradient{\QZ}
   \end{pmatrix}
}
Fluid motion can be described by the compressible Navier Stokes equations.
We follow the description in~\cite{dumbser2010arbitrary} for the Navier Stokes part.
The coupling with the advection-diffusion-reaction equation follows~\cite{hidalgo2011ader} which describes this equation set for the one-dimensional case.
In the following, we will denote the spatial coordinates as $\bm{x} = \left( x,z \right)$ and $\bm{x} = \left( x,y,z \right)$ for the two and three-dimensional case respectively.
The variable $t$ corresponds to the current physical time.
We solve the \pde{} in the conservative form% \pcref{eq:conservation-law-gradient}
\begin{equation}
 \label{eq:equation-set} 
 \begin{array}{l}
 \text{mass cons.} \\
 \text{momentum cons.} \\
 \text{energy cons.} \\
 \text{cont.\ gas} 
\end{array}
:
\quad
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE\\
    \QZ
    \end{pmatrix}}_{\Q}
  +
  \divergence{
  \underbrace{
  \left(
   \underbrace{\hyperFluxDef}_{\hyperFlux(\Q)}
+
\underbrace{\viscFluxDef}_{\viscFlux(\Q, \gradQ)}
  \right)}_{\flux(\Q, \gradQ)}}
 =
  \underbrace{
  \begin{pmatrix}
    \source[\Qrho\phantom{\Qrho}]\\
    \source[\Qj]\\
    \source[\QE]\\
    \source[\QZ]
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\end{equation}
which describes the time evolution of variables $\Q$ with respect to a flux $\flux(\Q, \gradQ)$ and a source $\source(\Q)$.
The vector of conserved quantities is given by
\begin{equation}
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE, \QZ \right),
\end{equation}
where $\Qrho$ is the density, $\Qj$ is the two or three-dimensional momentum, $\QE$ the energy density and $\QZ$ is the mass fraction of the chemical reactant.
The chemical reactant in our case is unburnt gas.

@@

\newcommand{\reactionTimescale}{\tau}
\newcommand{\reactionTemperature}{T_{\text{ign}}}
The chemical reactive source term is
\begin{equation}\label{eq:source-reaction}
  \source[\QZ] = - K(T) \QZ.
\end{equation}
We use the simple reaction model
\begin{equation}\label{eq:reaction-model}
\begin{cases}
  - \frac{1}{\reactionTimescale} & T > \reactionTemperature\\
  0 & \text{otherwise}
\end{cases}  
\end{equation}
where $\reactionTimescale$ is the timescale of the reaction and $\reactionTemperature$ is the activation temperature~\cite{hidalgo2011ader,helzel2000modified}.

\subsection{Boundary conditions}
To close the system we need to impose boundary conditions.\todo{Adapt this chapter to implementation!}

@@

\sidetitle{No-Slip}
A standard wall boundary condition for viscous fluid is the no-slip condition, where we assume that the fluid has a velocity of zero near the wall.
\todo{Check if this is the correct physical description!}
We enfore this by setting
\begin{align}
  \label{eq:no-slip}
  \begin{split}
  \Qrho^o &= \Qrho^i, \\
  \Qj^o &= -\Qj^i, \\
  \QE^0 &= \QE^i,\\
  {(\nabla Q)}^o &= {(\nabla Q)}^i,
  \end{split}
\end{align}
where a superscript of $o$ and $i$ denotes the values outside and inside of the boundary respectively.

@@

\sidetitle{Corrector}
The final step of the \aderdg{}-scheme is the corrector.
We multiply the system \cref{eq:conservation-law-gradient} by a space test function $\stestfunction$ and integrate over the space-time control volume.
This results, yet again, at the weak formulation of the \pde{}
\begin{equation}
  \label{eq:weak-pde}
\intdt{\intdcell{
\stestfunction[][v,n] \pdv{\Q}{t}
}}
+
\intdt{\intdcell{
    \stestfunction[][v,n] \left( \div{F(\Q, \gradQ } \right)
}}
=
\intdt{\intdcell{
    \stestfunction[][v,n] S(\Q, \bm{x}, t)
}}.
\end{equation}
We now insert our discrete approximations and apply the basis expansions \cref{eq:cell-approx-space,eq:cell-approx-space-time} to the space-time predictor.
The first part can be integrated by parts in time with the convenient result that only the boundary terms remain because the basis is constant in time.
We then integrate the flux divergence by parts in space and insert the Riemann solver to connect with the neighbouring elements.
Finally, we need to solve
\newcommand{\massMatrixDef}{\intdcell{
  \stestfunction[][v,n] \sbasis[][v,n']
}}
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\massMatrixDef
\right)
(\ssolCoeff[t^{n+1}][v,n'] - \ssolCoeff[t^n][v,n'])
&+
\left(\intdt{\intdcellb{
      \stestfunction[][v,n] \Riemann(\stpredictor[-], \gradient{\stpredictor[-]}; \stpredictor[+], \gradient{\stpredictor[+]}) \cdot \normal
}}\right)
-\\
\left(\intdt{\intdcell{
    \gradient{\stestfunction[][v,n]} \cdot  \stflux(\spredictor, \gradient{\spredictor})
}}\right)
&=
\left(\intdt{\intdcell{
      \stestfunction[][v,n] \stsource(\spredictor)
}}\right)
\end{split}
\end{align}
for the coefficients $\ssolCoeff[\nextTime][v,n]$ of the new solution.
Here the superscripts $\pm$ denotes values that are extrapolated to the boundaries.
Again, the computational effort can be vastly reduced by collecting the integrals into matrices and precomputing as much as possible.
The first term is the mass tensor with entries
\newcommand{\massMatrix}[1][]{\bm{M}_{#1}}
\begin{equation}
  \label{eq:mass-matrix}
  \massMatrix[v n n'] = \massMatrixDef = \volume \underbrace{\quadWeight[n] \delta_{n n'}}_{\text{diagonal}}
\end{equation}
which is, due to the orthogonal basis \pcref{eq:basis-orthogonality}, diagonal for each variable $v$ and thus trivial to invert.
We can then insert the definition of our basis function into our equation and arrive at
\begin{align}
  \massMatrix \left( \ssolCoeff[t+1] - \ssolCoeff[t] \right) &=
  \Delta t \left( \bm{a} - \bm{b} + \bm{s} \right) 
  \intertext{with}
a_{v,n} &= \sfluxCoeff[][n'] \intdcell{
  \sbasis[][v,n'] \gradient{\sbasis[][v,n]}(t)
}
\\
\label{eq:corrector-surface}
b_{v,n} &= \intdcellb{
    \stestfunction[][v,n] \Riemann \left( \spredictor[-], \gradient{\spredictor[-]}; \spredictor[+], \gradient{\spredictor[+]} \right) \cdot \normal
} \\
s_{v,n} &=
\ssourceCoeff[][v,n'] \intdcell{
  \sbasis[][v,n] \sbasis[][v,n']
}
,
\end{align}
where we evaluated the time-integrals by \cref{eq:integration-by-substitution}.
All quantities are now time-averaged~\pcref{eq:time-average} and are represented in the space basis~\pcref{eq:cell-approx-space}.
Note that we evaluate the Riemann problem for the time-averaged space-predictor.
This is justified by the linearity of our numerical flux~\cite{dumbser2008unified,dumbser2010arbitrary}.
The surface integral \pcref{eq:corrector-surface} can be evaluated by summing over the contributions of the faces of each cell, which are computed by integrating the extrapolated space predictor with Gaussian quadrature.
Finally, we can solve \cref{eq:corrector} for the new unknows $\ssolCoeff[k+1]$
\begin{align}\label{eq:update-predictor}
\begin{split}
  %\bm{u}^{n+1}
  \ssolCoeff[k+1][v]
  &= \ssolCoeff[k][v] + \Delta t (\operatorname{update}) \\
  \operatorname{update} &= \massMatrix[v]^{-1} \left( \bm{a}_v - \bm{b}_v + \bm{s}_v \right)
\end{split}
\end{align}
which is clearly of the form of a one-step scheme.

@@

\section{MUSCL-Hancock Finite Volume Scheme}\label{sec:muscl}
\newcommand{\cellAvg}[1][i,j]{U_{#1}}
\newcommand{\sign}{\operatorname{sign}}
\newcommand{\minmod}{\operatorname{minmod}}
\newcommand{\slope}[2][i,j]{s^{#2}_{#1}}
\newcommand{\gradCellAvg}[1][i,j]{\gradient{\cellAvg[#1]}}
\newcommand{\fluxX}{\flux_x}
\newcommand{\fluxY}{\flux_y}
We discussed a modern, high-order \dg{} method in the previous chapter.
While this scheme is efficient, it can become unstable, especially in case of discontinuities.
A simple alternative is a finite volume scheme.
We use the \muscl{} method\todo{Cite original paper (van Leer)}.
Our discussion and notation follows the one in~\cite{toro2009riemann}.
Similar to~\cite{toro2009riemann} we only describe the two-dimensional case, an extension to three dimensions is straigthforward.
\todo{Source term!}

@@

\todo{Reread dmp description, fix formulas.}
To combat the Gibbs phenomenon, we can make use of the discrete maximum principle (\textsc{dmp}).
This criterion marks cells as troubled that have a numerical solution that is significantly higher or lower than the solutions of its neighbours.
\begin{align}
\min (u(y, t^{\curTime})) - \delta &\leq u^*(y, t^{\nextTime}) \leq \max (u(y, t^{\curTime})) + \delta
%
%
\intertext{with tolerance}
\delta &= \max \left(
  \varepsilon_0, \varepsilon \left(
         \max(u(y,t^{\curTime})) -
         \min(u(y,t^{\curTime}))
  \right)
  \right)
\end{align}
The parameters $\varepsilon$ and $\varepsilon_0$ can be set freely.
The former is controls by which factor the local jump can differ from its neighbourhood, the latter is needed for zero jumps~\cite{dumbser2016simple}.
Note that this criterion is suspectible for false positives, this is especially a problem in the presence of a source term.
\todo{Mention problems with source term?}

>>>>

@@

e conclude this chapter by a description of finite volume limiting and adaptive mesh refinement (\amr{}).
The limiting combines both numerical methods to achieve a stable, high-order method and the \amr{} allows us to perform high quality simulations with a smaller computational cost.

\section{Reactive Compressible Navier Stokes Equations}\label{sec:navier-stokes}
\newcommand{\diffCoeff}{\varepsilon}
\newcommand{\hyperFluxDef}{
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure) \\
    \Qj \QZZ
  \end{pmatrix}
}
\newcommand{\viscFluxDef}{
  \begin{pmatrix}
     -\diffCoeff \gradient{\Qrho}\\
     \stressT (\Q, \gradQ)  \\
     \Qv \cdot \stressT (\Q, \gradQ) - \kappa \gradient{T}\\
     -\diffCoeff \gradient{\QZ}
   \end{pmatrix}
}

Fluid motion can be described by the compressible Navier Stokes equations.
We follow the description in~\cite{dumbser2010arbitrary} for the Navier Stokes part.
The coupling with the advection-diffusion-reaction equation follows~\cite{hidalgo2011ader} which describes this equation set for the one-dimensional case.
In the following, we will denote the spatial coordinates as $\bm{x} = \left( x,z \right)$ and $\bm{x} = \left( x,y,z \right)$ for the two and three-dimensional case respectively.
The variable $t$ corresponds to the current physical time.
We solve the \pde{} in the conservative form% \pcref{eq:conservation-law-gradient}
\begin{equation}
 \label{eq:equation-set} 
 \begin{array}{l}
 \text{mass cons.} \\
 \text{momentum cons.} \\
 \text{energy cons.} \\
 \text{cont.\ gas} 
\end{array}
:
\quad
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE\\
    \QZ
    \end{pmatrix}}_{\Q}
  +
  \divergence{
  \underbrace{
  \left(
   \underbrace{\hyperFluxDef}_{\hyperFlux(\Q)}
+
\underbrace{\viscFluxDef}_{\viscFlux(\Q, \gradQ)}
  \right)}_{\flux(\Q, \gradQ)}}
 =
  \underbrace{
  \begin{pmatrix}
    \source[\Qrho\phantom{\Qrho}]\\
    \source[\Qj]\\
    \source[\QE]\\
    \source[\QZ]
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\end{equation}
which describes the time evolution of variables $\Q$ with respect to a flux $\flux(\Q, \gradQ)$ and a source $\source(\Q)$.
The vector of conserved quantities is given by
\begin{equation}
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE, \QZ \right),
\end{equation}
where $\Qrho$ is the density, $\Qj$ is the two or three-dimensional momentum, $\QE$ the energy density and $\QZ$ is the mass fraction of the chemical reactant.
The chemical reactant in our case is unburnt gas.

@@

The chemical reactive source term is
\begin{equation}\label{eq:source-reaction}
  \source[\QZ] = - K(T) \QZ.
\end{equation}
We use the simple reaction model
\begin{equation}\label{eq:reaction-model}
\begin{cases}
  - \frac{1}{\reactionTimescale} & T > \reactionTemperature\\
  0 & \text{otherwise}
\end{cases}  
\end{equation}
where $\reactionTimescale$ is the timescale of the reaction and $\reactionTemperature$ is the activation temperature~\cite{hidalgo2011ader,helzel2000modified}.
\subsection{Boundary conditions}
To close the system we need to impose boundary conditions.\todo{Adapt this chapter to implementation!}

@@

\sidetitle{No-Slip}
A standard wall boundary condition for viscous fluid is the no-slip condition, where we assume that the fluid has a velocity of zero near the wall.
\todo{Check if this is the correct physical description!}
We enfore this by setting
\begin{align}
  \label{eq:no-slip}
  \begin{split}
  \Qrho^o &= \Qrho^i, \\
  \Qj^o &= -\Qj^i, \\
  \QE^0 &= \QE^i,\\
  {(\nabla Q)}^o &= {(\nabla Q^o)},
  \end{split}
\end{align}
where a superscript of $o$ and $i$ denotes the values outside and inside of the boundary respectively.

@@

\sidetitle{Corrector}
The final step of the \aderdg{}-scheme is the corrector.
We multiply the system \cref{eq:conservation-law-gradient} by a space test function $\stestfunction$ and integrate over the space-time control volume.
This results, yet again, at the weak formulation of the \pde{}
  \label{eq:weak-pde}
\intdt{\intdcell{
\stestfunction[][v,n] \pdv{\Q}{t}
}}
+
\intdt{\intdcell{
    \stestfunction[][v,n] \left( \div{F(\Q, \gradQ } \right)
}}
=
\intdt{\intdcell{
    \stestfunction[][v,n] S(\Q, \bm{x}, t)
}}.
We now insert our discrete approximations and apply the basis expansions \cref{eq:cell-approx-space,eq:cell-approx-space-time} to the space-time predictor.
The first part can be integrated by parts in time with the convenient result that only the boundary terms remain because the basis is constant in time.
We then integrate the flux divergence by parts in space and insert the Riemann solver to connect with the neighbouring elements.
Finally, we need to solve\todo{Fix notation, is tensor not matrix}
\newcommand{\massMatrixDef}{\intdcell{
  \stestfunction[][v,n] \sbasis[][v,n']
}}
\label{eq:corrector}
\left(
\massMatrixDef
\right)
(\ssolCoeff[t^{n+1}][v,n'] - \ssolCoeff[t^n][v,n'])
&+
\left(\intdt{\intdcellb{
      \stestfunction[][v,n] \Riemann(\stpredictor[-], \gradient{\stpredictor[-]}; \stpredictor[+], \gradient{\stpredictor[+]}) \cdot \normal
}}\right)
-\\
\left(\intdt{\intdcell{
    \gradient{\stestfunction[][v,n]} \cdot  \stflux(\spredictor, \gradient{\spredictor})
}}\right)
&=
\left(\intdt{\intdcell{
      \stestfunction[][v,n] \stsource(\spredictor)
}}\right)
for the coefficients $\ssolCoeff[\nextTime][v,n]$ of the new solution.
Here the superscripts $\pm$ denotes values that are extrapolated to the boundaries.
Again, the computational effort can be vastly reduced by collecting the integrals into matrices and precomputing as much as possible.
The first term is the mass tensor with entries
\newcommand{\massMatrix}[1][]{\bm{M}_{#1}}
  \label{eq:mass-matrix}
  \massMatrix[v n n'] = \massMatrixDef = \volume \underbrace{\quadWeight[n] \delta_{n n'}}_{\text{diagonal}}
which is, due to the orthogonal basis \pcref{eq:basis-orthogonality}, diagonal for each variable $v$ and thus trivial to invert.
We can then insert the definition of our basis function into our equation and arrive at
  \massMatrix \left( \ssolCoeff[t+1] - \ssolCoeff[t] \right) &=
  \Delta t \left( \bm{a} - \bm{b} + \bm{s} \right) 
  \intertext{with}
a_{v,n} &= \sfluxCoeff[][n'] \intdcell{
  \sbasis[][v,n'] \gradient{\sbasis[][v,n]}(t)
}
\\
\label{eq:corrector-surface}
b_{v,n} &= \intdcellb{
    \stestfunction[][v,n] \Riemann \left( \spredictor[-], \gradient{\spredictor[-]}; \spredictor[+], \gradient{\spredictor[+]} \right) \cdot \normal
} \\
s_{v,n} &=
\ssourceCoeff[][v,n'] \intdcell{
  \sbasis[][v,n] \sbasis[][v,n']
}
,
where we evaluated the time-integrals by \cref{eq:integration-by-substitution}.
All quantities are now time-averaged~\pcref{eq:time-average} and are represented in the space basis~\pcref{eq:cell-approx-space}.
Note that we evaluate the Riemann problem for the time-averaged space-predictor.
This is justified by the linearity of our numerical flux~\cite{dumbser2008unified,dumbser2010arbitrary}.
The surface integral \pcref{eq:corrector-surface} can be evaluated by summing over the contributions of the faces of each cell, which are computed by integrating the extrapolated space predictor with Gaussian quadrature.
Finally, we can solve \cref{eq:corrector} for the new unknows $\ssolCoeff[k+1]$
\begin{align}\label{eq:update-predictor}
  %\bm{u}^{n+1}
  \ssolCoeff[k+1][v]
  &= \ssolCoeff[k][v] + \Delta t (\operatorname{update}) \\
  \operatorname{update} &= \massMatrix[v]^{-1} \left( \bm{a}_v - \bm{b}_v + \bm{s}_v \right)
which is clearly of the form of a one-step scheme.

@@

\section{MUSCL-Hancock Finite Volume Scheme}\label{sec:muscl}
\todo{Mention dubious treatment of viscous boundary terms. Or later in exahype chapter.}
\newcommand{\cellAvg}[1][i,j]{U_{#1}}
\newcommand{\sign}{\operatorname{sign}}
\newcommand{\minmod}{\operatorname{minmod}}
\newcommand{\slope}[2][i,j]{s^{#2}_{#1}}
\newcommand{\gradCellAvg}[1][i,j]{\gradient{\cellAvg[#1]}}
\newcommand{\fluxX}{\flux_x}
\newcommand{\fluxY}{\flux_y}
We discussed a modern, high-order \dg{} method in the previous chapter.
While this scheme is efficient, it can become unstable, especially in case of discontinuities.
A simple alternative is a finite volume scheme.
We use the \muscl{} method\todo{Cite original paper (van Leer)}.
Our discussion and notation follows the one in~\cite{toro2009riemann}.
Similar to~\cite{toro2009riemann} we only describe the two-dimensional case, an extension to three dimensions is straigthforward.
\todo{Source term!}

@@

\todo{Reread dmp description, fix formulas.}
To combat the Gibbs phenomenon, we can make use of the discrete maximum principle (\textsc{dmp}).
This criterion marks cells as troubled that have a numerical solution that is significantly higher or lower than the solutions of its neighbours.
\begin{align}\label{eq:limiting-dmp}
\min (u(y, t^{\curTime})) - \delta &\leq u^*(y, t^{\nextTime}) \leq \max (u(y, t^{\curTime})) + \delta
%
%
\intertext{with tolerance}
\delta &= \max \left(
  \varepsilon_0, \varepsilon \left(
         \max(u(y,t^{\curTime})) -
         \min(u(y,t^{\curTime}))
  \right)
  \right)
\end{align}
The parameters $\varepsilon$ and $\varepsilon_0$ can be set freely.
The former is controls by which factor the local jump can differ from its neighbourhood, the latter is needed for zero jumps~\cite{dumbser2016simple}.
Note that this criterion is suspectible for false positives, this is especially a problem in the presence of a source term.
\todo{Mention problems with source term?}
